Modelling the City of Melbourne Pedestrian Data

### 2017-05-07 presentation/proposal_pres.Rmd
majority of content complete - just needs some polishing up and potentially some trimming

```{r setup, include=FALSE, echo = FALSE, warning=FALSE, message=FALSE}
options(htmltools.dir.version = FALSE, scipen = 100)
library(readr)
library(dplyr)
library(ggplot2)
library(ggmap)
library(knitr)
library(pander)
library(lubridate)
  ped_df <- read_csv("../data/Pedestrian_volume__updated_monthly_.csv")

@@

- Due to the multiple seasonal patterns in the time series, typical models like AR, ARMA and ARIMA do not work very well
  + Modelling using other sources of data is being explored, in particular weather data
  + Alternative models, such as a Multiple Seasonal (MS) process available through the `forecast` R package, should also be investigated

@@

```{r, echo = FALSE, message=FALSE, out.height="500px", fig.retina=2}
  ggplot(ped_df[ped_df$Sensor_ID == "3", ]) + 
    geom_smooth(aes(x = Timeplot, y = Hourly_Counts, colour = Month), se = F) + 
    xlab("Time of the Day") + ylab("Pedestrian Count") + 
    ggtitle(paste("Monthly Average Hourly Counts at", ped_loc$`Sensor Description`[ped_loc$`Sensor ID` == "3"], "by Day")) +
    facet_wrap(~ Day, nrow = 2) + theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom") +
    scale_x_continuous(limits = c(0,24))
```

>>>>

```{r setup, include=FALSE, echo = FALSE, warning=FALSE, message=FALSE}
options(htmltools.dir.version = FALSE, scipen = 100)
library(readr)
library(dplyr)
library(ggplot2)
library(ggmap)
library(knitr)
library(lubridate)
  ped_df <- read_csv("../data/Pedestrian_volume__updated_monthly_.csv")

@@

- Due to the multiple seasonal patterns in the time series, typical models like AR, ARMA and ARIMA do not work very well
  + Modelling using other sources of data is being explored, in particular weather data
  + Alternative models, such as a Multiple Seasonal (MS) process available through the `forecast` R package, should also be investigated
  + Machine learning techniques, such as clustering and other non-parametric estimation methods, will also be considered

@@

```{r, echo = FALSE, message=FALSE, out.height="480px", fig.retina=2, fig.width = 12}
  ggplot(ped_df[ped_df$Sensor_ID == "3", ]) + 
    geom_smooth(aes(x = Timeplot, y = Hourly_Counts, colour = Month), se = F) + 
    xlab("Time of the Day") + ylab("Pedestrian Count") + 
    ggtitle(paste("Monthly Average Hourly Counts at", ped_loc$`Sensor Description`[ped_loc$`Sensor ID` == "3"], "by Day")) +
    facet_wrap(~ Day, nrow = 2) + theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom") +
    scale_x_continuous(limits = c(0,24))
```

```{r, echo = FALSE, message=FALSE, out.height="480px", fig.retina=2, fig.width = 12}
  ggplot(ped_df[ped_df$Sensor_ID == "3", ]) + 
    geom_density(aes(x = Hourly_Counts, colour = Day), adjust = 2, alpha = 0.2) + 
    xlab("Time of the Day") + ylab("Density") + 
    ggtitle(paste("Kernel Density Estimates of Hourly Counts at", ped_loc$`Sensor Description`[ped_loc$`Sensor ID` == "3"], "by Day")) +
    theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom")
```

---
class: middle

## Visualisation: importance of design

As we can see, there are many different ways to present the same data

When designing visualisations for pedestrian data, we need to take into consideration:
- Preattentive processing
- Change blindness

Need to communicate the intended message to the person viewing the plots

## Interactive Visualisation: Shiny apps

The package `shiny` allows us to build web apps using R's capabilities in the background
We can add interactive inputs to visualisations.

An example of what we can make:

class: inverse, middle

# Research Plan

### The plan to approach this research project is as follows:

- Imputation that incorporates time and space
- Creating a predictive model for each sensor
- Investigation of visual methods for space-time counts
- Development of a Shiny app: interface design, computation and data backend
- Generation of a dynamic spatial model (if time permits)

class: inverse, middle, center

# Questions or comments?

# Acknowledgements

##### Thanks to Di Cook and Earo Wang for helping me thus far in the project!

This presentation was written in R Markdown using the template `xaringan`.
Other packages used for this presentation are: `readr, dplyr, ggplot2, ggmap, knitr, lubridate`
